---
title: "Assignment1"
author: "Esmma"
date: '2022-09-27'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(dplyr)
library(nycflights13)
library(ggplot2)
```

```{r}
#How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
sum(is.na(flights$dep_time))
colSums(is.na(flights))
```

8255 flights have a missing department time. Other variables that are missing are dep_delay, arr_time, arr_delay, tailnum, and air_time. These rows represent a canceled flight.

```{r}
#Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
flights_times <- mutate(flights,
  dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
  sched_dep_time_mins = (sched_dep_time %/% 100 * 60 +
    sched_dep_time %% 100) %% 1440
)
select(
  flights_times, dep_time, dep_time_mins, sched_dep_time,
  sched_dep_time_mins
)
head(flights_times)
```

```{r}
#Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()
canceled_flights <- flights %>% 
  group_by(year,month,day) %>% 
  mutate(canceled = if_else(
    condition = is.na(dep_time),
    true = T,
    false = F,
    missing = T
  )) %>% 
  summarise(total_flights = n(), missing_flights = sum(canceled))

ggplot(canceled_flights) +
  geom_point(aes(x = total_flights, y = missing_flights)) 
```

```{r}
cancellations_and_delays <- flights %>% 
  group_by(year,month,day) %>% 
  mutate(canceled = if_else(
    condition = is.na(dep_time),
    true = T,
    false = F,
    missing = T
  )) %>% 
  summarise(total = n(), missing = sum(canceled), prop_cancel = mean(canceled),avg_arrdelay = mean(arr_delay, na.rm = T), avg_depdelay = mean(dep_delay,na.rm = T)) %>% 
  ungroup() 

ggplot(cancellations_and_delays) +
  geom_point(aes(x = avg_depdelay, y = prop_cancel ))

ggplot(cancellations_and_delays) +
  geom_point(aes(x = avg_arrdelay, y = prop_cancel ))
```

There is a strong increasing relationship between proportion of cancellations and average departure delay and a strong increasing relationship between proportion of cancellations and average arrival delay.



